SIMD\slash SIMT-vectorized Sparse Chemical Kinetic Jacobian\slash Thermo-chemical Source Term Evaluation

### 2018-02-18 paper.tex
updates

\newcommand{\pluseq}{\mathrel{{+}{=}}}
\newcommand{\minuseq}{\mathrel{{-}{=}}}
\newcommand{\ns}{N_{sp}}
\newcommand{\nr}{N_{reac}}
\newcommand{\conp}{CONP}
\newcommand{\conv}{CONV}
\newcommand{\dconp}{\ifmmode\text{for \conp,}\else for \conp,\fi}
\newcommand{\dconv}{\ifmmode\text{for \conv,}\else for \conv.\fi}
\newcommand{\Ru}{\mathcal{R}}

@@

\begin{table}[htb]
\centering
\begin{tabular}{@{}l l l@{}}
\toprule
CPU Model        & Xeon X5650      & E5-2690 V3     \\
\midrule
Instruction Set  & SSE4.2 	   & AVX2 	    \\
Vector Width     & \SI{128}{\bit}  & \SI{256}{\bit} \\
Cores            & $2 \times 6$    & $2 \times 12$  \\
Identifier       & \texttt{sse4.2} & \texttt{avx2}  \\
OpenCL Version   & \num{1.2}       & \num{1.2}      \\
\bottomrule
\end{tabular}
\caption{The Intel CPU configurations used in this study. The identifier field will be used as a shorthand descriptor in the performance plots to quickly identify the CPU type.}
\label{t:cpus}
\end{table}

\begin{table}
\centering
\begin{tabular}{@{}l l l l l@{}}
\toprule
NVIDIA Model   & Tesla C2075    & Tesla K40m    \\
\midrule
Driver Version & \num{384.81}   & \num{387.26}  \\
CUDA Cores     & \num{448}      & \num{2880}    \\
Identifier     & \texttt{C2075} & \texttt{K40m} \\
OpenCL Version & \num{1.1}	& \num{1.2}	\\
\bottomrule
\end{tabular}
\caption{The NVIDIA GPU configurations used in this study.  NVIDIA's OpenCL runtime is provided with the graphic driver, rather than any specific version of CUDA.
Again, the identifier field will be used as a shorthand descriptor during analysis of the performance results.}
\label{t:gpus}
\end{table}

@@

\subsection{Performance}
\label{S:results}
The performance studies in this work were run on the platforms listed in~\cref{t:cpus,t:gpus}.
Runtimes in each case were averaged over ten runs, each using the same set of PaSR conditions utilized in validation.
The wrapping code---responsible for initializing\slash transfering memory, reading input, etc.---was compiled with \texttt{gcc v5.4.0} on the \texttt{avx2}\slash\texttt{K40m} platforms, and \texttt{gcc 4.8.5} on the \texttt{sse4.2}\slash\texttt{C2075} machines.
The optimization level ``\texttt{-O3}'' was used in all cases, and no ``fast-math'' OpenCL optimizations were enabled.
Additionally, the ``exact'' form of the Jacobian (as opposed the the ``appropriate'' form discussed in~\cref{S:sparsity}) was used in all cases.

\Cref{f:source} shows the runtime of \texttt{pyJac} on the test platforms.

\begin{figure}[htb]
   \centering
  \begin{subfigure}{0.45\linewidth}
      \includegraphics[width=\textwidth]{intel_source.pdf}
      \caption{The ``exact'' Jacobian.}
      \label{F:intel_source_noorm}
  \end{subfigure}
 \label{f:source}
\end{figure}

>>>>

\newcommand{\pluseq}{\mathrel{{+}{=}}}
\newcommand{\minuseq}{\mathrel{{-}{=}}}
\newcommand{\ns}{N_{sp}}
\newcommand{\nr}{N_{reac}}
\newcommand{\conp}{CONP}
\newcommand{\conv}{CONV}
\newcommand{\dconp}{\ifmmode\text{for \conp,}\else for \conp,\fi}
\newcommand{\dconv}{\ifmmode\text{for \conv,}\else for \conv.\fi}
\newcommand{\Ru}{\mathcal{R}}
\newcommand{\avx}{\texttt{avx2}}
\newcommand{\sse}{\texttt{sse4.2}}
\newcommand{\gpuold}{\texttt{C2075}}
\newcommand{\gpunew}{\texttt{K40m}}

@@

\begin{table}[htb]
\centering
\begin{tabular}{@{}l l l@{}}
\toprule
CPU Model        & Xeon X5650      & E5-2690 V3     \\
\midrule
Instruction Set  & SSE4.2 	   & AVX2 	    \\
Vector Width     & \SI{128}{\bit}  & \SI{256}{\bit} \\
Cores            & $2 \times 6$    & $2 \times 12$  \\
Identifier       & \sse & \avx  \\
OpenCL Version   & \num{1.2}       & \num{1.2}      \\
\bottomrule
\end{tabular}
\caption{The Intel CPU configurations used in this study. The identifier field will be used as a shorthand descriptor in the performance plots to quickly identify the CPU type.}
\label{t:cpus}
\end{table}

\begin{table}
\centering
\begin{tabular}{@{}l l l l l@{}}
\toprule
NVIDIA Model   & Tesla C2075    & Tesla K40m    \\
\midrule
Driver Version & \num{384.81}   & \num{387.26}  \\
CUDA Cores     & \num{448}      & \num{2880}    \\
Identifier     & \gpuold & \gpunew \\
OpenCL Version & \num{1.1}	& \num{1.2}	\\
\bottomrule
\end{tabular}
\caption{The NVIDIA GPU configurations used in this study.  NVIDIA's OpenCL runtime is provided with the graphic driver, rather than any specific version of CUDA.
Again, the identifier field will be used as a shorthand descriptor during analysis of the performance results.}
\label{t:gpus}
\end{table}

@@

\subsection{Performance}
\label{S:results}
The performance studies in this work were run on the platforms listed in~\cref{t:cpus,t:gpus}.
Runtimes in each case were averaged over ten runs, each using the same set of PaSR conditions utilized in validation.
The wrapping code---responsible for initializing\slash transfering memory, reading input, etc.---was compiled with \texttt{gcc v5.4.0} on the \avx\slash\gpunew platforms, and \texttt{gcc 4.8.5} on the \sse\slash\gpuold machines.
The optimization level ``\texttt{-O3}'' was used in all cases, and no ``fast-math'' OpenCL optimizations were enabled.
Additionally, the ``exact'' form of the Jacobian (as opposed the the ``appropriate'' form discussed in~\cref{S:sparsity}) was used in all cases.

\Cref{f:source} explores the performance of the source term evaluations generated by \texttt{pyJac} on the CPU test platforms listed in~\cref{t:cpus}.
In~\cref{F:intel_source_noorm}, the mean runtime per initial condition is shown for both the~\avx and~\sse CPUs, using both Intel OpenCL and OpenMP.
This normalization is chosen to account for the varying numbers of conditions in the PaSR databases for each model (\cref{T:error}}).
For both CPUs, the OpenMP implementation is the slowest over all models; interestingly the parallel (non-vectorized) Intel OpenCL code is slightly faster than OpenMP in all cases.\todo{rerun OpenMP w/ mtune=native}
As expected, the \avx machine is moderately faster than the \sse CPU; additionally, the wide-vectorized OpenCL code is significantly faster than either the OpenMP or unvectorized OpenCL codes on both processors.
In~\cref{F:intel_source}, the extent of this speedup is detailed; we note that the speedup displayed is normalized per-machine, such that the speedup of the wide-vectorized OpenCL code can be readily assessed.
On both machines, the unvectorized OpenCL code is \textasciitilde\SIrange{1.25}{1.5}{$\times$} faster than the baseline parallel OpenMP code.
Additionally, the wide-vectorized OpenCL code is \textasciitilde\SIrange{2.5}{3}{$\times$} and \textasciitilde\SIrange{3.5}{4}{$\times$} faster than the OpenMP code for the \sse and \avx machines respectively.
In contrast,~\cref{F:pocl_source} shows the runtime of deep, wide and unvectorized OpenCL codes on the POCL runtime, as compared to OpenMP parallelization.
No speedup is achieved for either vectorization type on either CPU; indeed the OpenMP is faster in both cases.
The reasons for this lack of vectorization are quite technical and somewhat outside the scope of this work, however briefly speaking it is possible vectorization might be achieved on POCL via combination of relaxed math (``-cl-fast-relaxed-math''), use of explicit OpenCL vector types and future improvements to LLVM's vectorization module.
This will be explored more in future work, but for the moment POCL is still quite useful as a validation tool.

\begin{figure}[htb]
   \centering
  \begin{subfigure}{0.45\linewidth}
      \includegraphics[width=\textwidth]{intel_source_nonorm.pdf}
      \caption{The mean runtime per condition for each chemical model using the Intel OpenCL and OpenMP on both CPUs studied; the names in the legend correspond to the identifiers listed in~\cref{t:cpus}}
      \label{F:intel_source_noorm}
  \end{subfigure}
  \begin{subfigure}{0.45\linewidth}
      \includegraphics[width=\textwidth]{intel_source.pdf}
      \caption{The speedup achieved over the baseline OpenMP parallelization by both the unvectorized and wide-vectorized Intel OpenCL codes for each chemical model.}
      \label{F:intel_source}
  \end{subfigure}
  \\
  \begin{subfigure}{0.45\linewidth}
      \includegraphics[width=\textwidth]{pocl_source_nonorm.pdf}
      \caption{The mean runtime per condition of the Portable OpenCL runtime compared to OpenMP parallelization for each chemical model.}
      \label{F:pocl_source}
  \end{subfigure}
 \label{f:source}
\end{figure}
